The importance of transient resonances in extreme-mass-ratio inspirals

### 2016-08-06 asymptotic.tex
Jumps done

Having constructed the solution, we see that the lowest-order evolution corresponds to the adiabatic solution; the deviations come in at the following order. When we switch from the pre-resonance solution to the post-resonance solution, there is a change in the sign of $\widehat{\lambda}$. Therefore, when matching the post-resonance solution $\widehat{\varrho}_{1/2}$ and $\widehat{\kappa}_{1/2}$ also change sign: there is a change of
\begin{align}
\Delta q = {} & 2 \eta^{1/2}\widehat{\kappa}_{1/2}, \\
\Delta J = {} & 2 \eta^{1/2}\widehat{\varrho}_{1/2}
\label{eq:jumps}
\end{align}
across the resonance~\cite{Kevorkian1987}. We are not particularly interested in the deviation in $J$, of greater concern is the change in the orbital parameters $\{E,L_z,Q\}$. Assuming that there is a smooth transformation that maps between $J$ and these, then, to lowest order, we can calculate the deviation relative to the adiabatic prescription by substituting the forcing functions $G^{(1)} \rightarrow G_a^{(1)}$, where $G_a^{(1)}$ describes the evolution of $\mathcal{I}_a$ through the effects of the self-force. This result is quoted by Flanagan and Hinderer~\cite{Flanagan2012}. The change in the orbital parameters is determined by the forcing functions, hence it is essential to have an accurate self-force model.

As a final step in understanding our result, we switch from Mino time to coordinate time. An appropriate redefinition of the forcing functions can be done by scaling by $\Gamma$, we define
\begin{equation}
F_a^{(1)} = \dfrac{G_a^{(1)}}{\Gamma},
\end{equation}
such that the equation of motion becomes
\begin{equation}
\left\langle \diff{\mathcal{I}_a}{t}\right\rangle_{q'} =  \eta\sum_s F_{a,\,s}^{(1)}(\boldsymbol{\mathcal{I}})\exp(is q) + \order{\eta^2}.
\end{equation}
Here we have made the averaging over $q'$ explicit to show that the equation is only defined as an orbital average: not only does our asymptotic expansion average out oscillations over an orbit in $q'$, but in converting from $\lambda$ to $t$ we have used $\Gamma$ which is an orbital average.
From \eqnref{omega-series}, we recognise that
\begin{equation}
\varpi_1 = \partialdiff{\omega}{\widetilde{\lambda}} = \dfrac{\Gamma^2}{\eta}\left\langle\dot{\Omega}\right\rangle_{q'}.
\end{equation}
We have used the averaged form of $\dot{\Omega}(t)$ as this is appropriate. Using these to adapt equations (\ref{eq:J-1/2}) and (\ref{eq:jumps}), we obtain
\begin{align}
\Delta \mathcal{I}_a = {} & \eta\sum_{s\,\neq\,0}F_{a,\,s}^{(1)}(\boldsymbol{\mathcal{I}}_\star)\left[\dfrac{2\pi}{\left|s \left\langle\dot{\Omega}\right\rangle_{q'}\right|}\right]^{1/2} \nonumber \\*
 {} & \times \left.\exp\left[i\left(s \widehat{\kappa}_0 + \dfrac{\pi}{4}\sgn s\dot{\Omega}\right)\right]\right. \\
 = {} & \eta\sum_{s\,\neq\,0}F_{a,\,s}^{(1)}(\boldsymbol{\mathcal{I}}_\star)\tau_{\mathrm{res},\,s}\exp\left[i\left(s \widehat{\kappa}_0 + \dfrac{\pi}{4} \sgn s\dot{\Omega} \right)\right],
\end{align}
using \eqnref{T-res-s} and representing the values on resonance of $E$, $L_z$ and $Q$ with $\boldsymbol{\mathcal{I}}_\star$.

>>>>

Having constructed the solution, we see that the lowest-order evolution corresponds to the adiabatic solution; the deviations come in at the following order. When we switch from the pre-resonance solution to the post-resonance solution, there is a change in the sign of $\widehat{\lambda}$. Therefore, when matching the post-resonance solution $\widehat{\varrho}_{1/2}$ and $\widehat{\kappa}_{1/2}$ also change sign: there is a change of
\begin{align}
\Delta q = {} & 2 \eta^{1/2}\widehat{\kappa}_{1/2}, \\
\Delta J = {} & 2 \eta^{1/2}\widehat{\varrho}_{1/2}
\label{eq:jumps}
\end{align}
across the resonance~\cite{Kevorkian1987}. We are not particularly interested in the deviation in $J$, of greater concern is the change in the orbital parameters $\{E,L_z,Q\}$. Assuming that there is a smooth transformation that maps between $J$ and these, then, to lowest order, we can calculate the deviation relative to the adiabatic prescription by substituting the forcing functions $G^{(1)} \rightarrow G_a^{(1)}$, where $G_a^{(1)}$ describes the evolution of $\mathcal{I}^a$ through the effects of the self-force. This result is quoted by Flanagan and Hinderer~\cite{Flanagan2012}. The change in the orbital parameters is determined by the forcing functions, hence it is essential to have an accurate self-force model.

As a final step in understanding our result, we switch from Mino time to coordinate time. An appropriate redefinition of the forcing functions can be done by scaling by $\Gamma$, we define
\begin{equation}
F_a^{(1)} = \dfrac{G_a^{(1)}}{\Gamma},
\end{equation}
such that the equation of motion becomes
\begin{equation}
\left\langle \diff{\mathcal{I}^a}{t}\right\rangle_{q'} =  \eta\sum_s F_{a,\,s}^{(1)}(\boldsymbol{\mathcal{I}})\exp(is q) + \order{\eta^2}.
\end{equation}
Here we have made the averaging over $q'$ explicit to show that the equation is only defined as an orbital average: not only does our asymptotic expansion average out oscillations over an orbit in $q'$, but in converting from $\lambda$ to $t$ we have used $\Gamma$ which is an orbital average.
From \eqnref{omega-series}, we recognise that
\begin{equation}
\varpi_1 = \partialdiff{\omega}{\widetilde{\lambda}} = \dfrac{\Gamma^2}{\eta}\left\langle\dot{\Omega}\right\rangle_{q'}.
\end{equation}
We have used the averaged form of $\dot{\Omega}(t)$ as this is appropriate. Using these to adapt equations (\ref{eq:J-1/2}) and (\ref{eq:jumps}), we obtain
\begin{align}
\Delta \mathcal{I}^a = {} & \eta\sum_{s\,\neq\,0}F_{a,\,s}^{(1)}(\boldsymbol{\mathcal{I}}_\star)\left[\dfrac{2\pi}{\left|s \left\langle\dot{\Omega}\right\rangle_{q'}\right|}\right]^{1/2} \nonumber \\*
 {} & \times \left.\exp\left[i\left(s \widehat{\kappa}_0 + \dfrac{\pi}{4}\sgn s\dot{\Omega}\right)\right]\right. \\
 = {} & \eta\sum_{s\,\neq\,0}F_{a,\,s}^{(1)}(\boldsymbol{\mathcal{I}}_\star)\tau_{\mathrm{res},\,s}\exp\left[i\left(s \widehat{\kappa}_0 + \dfrac{\pi}{4} \sgn s\dot{\Omega} \right)\right],
\end{align}
using \eqnref{T-res-s} and representing the values on resonance of $E$, $L_z$ and $Q$ with $\boldsymbol{\mathcal{I}}_\star$.

### 2016-08-06 cut_figures.tex
Jumps done

>>>>

For each system, we compute the overlap between the full evolution and the family of adiabatics given by our set of match times $t\sub{match}$ described in the previous section. The largest overlap in each case, plotted in \figref{overlap-vs-q}, appears to be roughly independent of the value of $q$ and hence with the relative sizes of the orbital parameter jumps.
\begin{figure}
\centering
\includegraphics[width=0.46\textwidth]{overlap_vs_q}
\caption{\label{fig:overlap-vs-q}The overlap between the full and adiabatic evolutions, maximised over the family of generated adiabatics, as a function of the extracted phase parameter $q$. Each dot represents a system with our illustrative parameters and a random choice of the initial radial phase. The red line indicates the expected overlap assuming that the waveforms match exactly post-resonance, but have zero overlap pre-resonance.}
\end{figure}
The best performing adiabatic for the majority of our systems matched at the end of the inspiral. We can therefore estimate the value of the overlap by assuming that the adiabatic waveforms recover the entire SNR in the region post-resonance, but none at all in the region pre-resonance. We expect the overlap to scale with the fraction of time spent post-resonance, the value of which is demonstrated by the red line in \figref{overlap-vs-q}. There is not an exact equivalence because of the frequency-dependent noise of eLISA; cycles near the end of the inspiral accumulate proportionally more SNR at a frequency closer to the minimum of the eLISA noise curve.
Systems with $|q| \approx \pi/2$ (corresponding to small jumps in $E$ and $L_z$) have higher than average overlaps due to additional accumulation in the pre-resonance region. The adiabatic models of these systems dephase more slowly, but still do not produce very high overlaps due to the jump in $Q$. A small fraction of systems have a lower overlap than expected, at around $0.5$. This is due to a sub-optimal choice of the matching times; making small adjustments to $t\sub{match}$ can lead to higher overlaps, but the exact adjustments required are difficult to predict in advance.
